%\documentclass[12pt]{scrreprt}
\documentclass[12pt]{report} 

% language may be romanian or english (default is english)
% type may be bachelor or master (default is bachelor)
\usepackage[language=english, type=bachelor]{style}
\usepackage{amsmath}
\usepackage{adjustbox}


%\geometry{a4paper,top=2.5cm,left=3cm,right=2.5cm,bottom=2.5cm}
%in style
%controlling the appearance of your headers and footers
\usepackage{fancyhdr}
\pagestyle{fancy}
\lhead{}
\chead{}
\renewcommand{\headrulewidth}{0.2pt}
\renewcommand{\footrulewidth}{0.2pt}

\begin{document}

\specialization{COMPUTER SCIENCE ROMANIAN}	
\title{Detecting Signs of Depression in a Cross-Linguistic Context Using Machine Learning}					   
\author{Dinică Mircea}											
\supervisor{Lect. PhD. Lupea Mihaiela}				
				
\maketitle


\newpage
\thispagestyle{empty}
\mbox{}
\newpage
\pagenumbering{roman} 

\cleardoublepage
ABSTRACT
\vspace{0.5cm}	
\hrule
\vspace{0.5cm}	
%\cleardoublepage

The purpose of this paper is to develop Artifical Intelligence (AI) classifiers that detect if the text shows signs of depression or not in two languages: English and Romanian. Depression is a major problem right now, which if not addressed can cause serious problems. With the help of NLP and AI, it is desired to develop a tool for giving an opportunity to its users to have an easy way to get an initial assessment of their texts, in cases such as private text messages, their own thoughts or even help professionals reach possible clients in order to help them. 

The two classifiers will be created using the same dataset for training. The model that processes English text will use the original dataset in English. For the Romanian model, the dataset is translated using a Python library with the Google Translate API. For pre-processing, a tool called Linguistic Inquiry and Word Count will be used. The English model will use the latest version, LIWC-22, developed in 2022. The Romanian model will use the most recent translated version, LIWC-2015, developed in 2015. Both classifiers will use the Random Forest model.

As a practical tool for the algorithm, a chrome extension which expects as input a text was developed. The output is the class which the model predicted with a corresponding message advising the user to investigate further with the author of the text. This will be computed on a web back-end server, where the AI classifier is used, respecting the REST architectural style. The chosen technologies for developing the tool are Python with Flask framework for the back-end server and React, the JavaScript library for the front-end server.

The application can work in multiple languages because it uses the same Python library that was used to translate the dataset. The input text is translated when the language it is written in is different from Romanian or English.  

\tableofcontents


\newpage
\pagenumbering{arabic}

\input{chapters/chapter1_introduction}
%\addcontentsline{toc}{chapter}{Introducere}
%\addcontentsline{toc}{chapter}{Introduction}

\input{LaTeX Bachelor Thesis Depression Signs Detection/chapters/chapter2_theoretical_background}
\input{chapters/chapter3_dataset_and_LIWC}
\input{chapters/chapter4_model}
\input{LaTeX Bachelor Thesis Depression Signs Detection/chapters/chapter5_practicApp}
\input{chapters/chapter6_conclusions}
%\addcontentsline{toc}{chapter}{Concluzii}
%\addcontentsline{toc}{chapter}{Conclusions}

\bibliography{references}

\end{document}
%TODO: cand cumpar licenta pt liwc schimbat inloc de 64 input attributes cu cate o sa am